



We provide a quantum sample algorithm for the Uhlmann transformation in the mixed sample access model.
The key idea for construction is to use a property of the maximally entangled state $\ket{\Phi}^{\sA\sB} = \f{1}{\sqrt{d_\sA}}\sum_{i=1}^{d_\sA}\ket{i}^\sA\ket{i}^{\sB}$: for any state $\omega^\sA$,
\begin{equation}
(\sqrt{\omega^\sA} \otimes \bI^{\sB})\ket{\Phi}^{\sA\sB}
= \f{1}{\sqrt{d_\sA}}\ket{\omega_{\rm c}}^{\sA\sB},
\end{equation}
where $\ket{\omega_{\rm c}}^{\sA\sB}$ is the canonical purified state of $\omega^\sA$. 
Note that $\tr_\sB[\ketbra{\omega_{\rm c}}{\omega_{\rm c}}^{\sA\sB}] = \omega^\sA$.
It is natural to assume that $d_\sA = d_\sB$ in this model.
Our approach is to first prepare canonical purified states $\ket{\rho_{\rm c}}^{\sA\sB}$ and $\ket{\sigma_{\rm c}}^{\sA\sB}$ from multiple copies of given states $\rho^{\hat{\sA}}$ and $\sigma^{\hat{\sA}}$, and then to perform the Uhlmann transformation algorithm for the purified sample access model described in the previous section.  
This is to find the Uhlmann transformation from $\ket{\rho_{\rm c}}^{\sA\sB}$ to $\ket{\sigma_{\rm c}}^{\sA\sB}$.


The statement is as follows, where the system $\sH$ is a two-qubit system. 
Remember that $s_{\rm min}$ and $r$ are the minimum non-zero singular value and the rank of $\sqrt{\sigma^\sA}\sqrt{\rho^\sA}$, respectively. Also, $\rho_{\rm min}$ and $\sigma_{\rm min}$ are the minimum non-zero eigenvalues of $\rho^\sA$ and $\sigma^\sA$, respectively.





\begin{theorem}[Uhlmann transformation algorithm in the mixed sample access model]
\label{thm:Uhlmann mixed state sample}
Let $\delta \in (0, 1)$ and $\chi \in \{\diamond, \tF\}$. 
Then, the quantum sample algorithm $\mathtt{UhlmannMixedSample}$ given by \Cref{alg:Uhl mixed sample} satisfies the following.

A quantum channel $\cJ_\diamond^{\sB\sH\hat{\sA}\hat{\sB}}$ given by $\cJ_\diamond^{\sB\sH\hat{\sA}\hat{\sB}}\circ\cP_{\til{\sigma}_{\rm c}}^{\bC\rarr\hat{\sA}\hat{\sB}}  
    =\mathtt{UhlmannMixedSample}(\rho^{\hat{\sA}}, \sigma^{\hat{\sA}}; \delta, \diamond)$,
with some state $\til{\sigma}_{\rm c}^{\hat{\sA}\hat{\sB}}$, satisfies that
\begin{equation}
    \f{1}{2}\big\|\cJ_\diamond^{\sB\sH\hat{\sA}\hat{\sB}}\circ\cP_{\ket{0}}^{\bC\rarr\sH}\circ\cP_{\til{\sigma}_{\rm c}}^{\bC\rarr\hat{\sA}\hat{\sB}} - \cU_{\rm ideal}^{\sB\sH\hat{\sA}\hat{\sB}}\circ\cP_{\ket{0}\ket{\sigma_{\rm c}}}^{\bC\rarr\sH\hat{\sA}\hat{\sB}}\big\|_\diamond \leq \delta,
\end{equation}
where $\til{U}_{\rm ideal}^{\sB\sH\hat{\sA}\hat{\sB}}$ is an exact block-encoding unitary of $\ketbra{\rho_{\rm c}}{\sigma_{\rm c}}^{\hat{\sA}\hat{\sB}} \otimes V^\sB$ and $V^\sB$ is the Uhlmann partial isometry.
The algorithm uses
\begin{align}
    \zeta_\diamond = \til{\cO}\Big(\f{d_\sA}{\delta^2 s_{\rm min}^3} \min\Big\{\f{1}{\rho_{\rm min}^2} + \f{1}{\sigma_{\rm min}^2}, \f{d_\sA^2}{\delta^4 s_{\rm min}^4}\Big\}\Big),
\end{align}
samples of $\rho^{\hat{\sA}}$ and $\sigma^{\hat{\sA}}$.

A quantum channel $\cJ_\tF^{\sB\sH\hat{\sA}\hat{\sB}}$ given by $\cJ_\tF^{\sB\sH\hat{\sA}\hat{\sB}}\circ\cP_{\til{\sigma}_{\rm c}}^{\bC\rarr\hat{\sA}\hat{\sB}} 
    = \mathtt{UhlmannMixedSample}(\rho^{\hat{\sA}}, \sigma^{\hat{\sA}}; \delta, \tF)$,
with some state $\til{\sigma}_{\rm c}^{\hat{\sA}\hat{\sB}}$, satisfies that
\begin{align}
    \rF\big(\cT^\sB(\ketbra{\rho_{\rm c}}{\rho_{\rm c}}^{\sA\sB}), \ket{\sigma_{\rm c}}^{\sA\sB}\big) \geq \rF(\rho^\sA, \sigma^\sA) - \delta,
\end{align}
where $\cT^{\sB}= \tr_{\sH\hat{\sA}\hat{\sB}}\circ\cJ_\tF^{\sB\sH\hat{\sA}\hat{\sB}}\circ\cP_{\ket{0}}^{\bC\rarr\sH}\circ\cP_{\til{\sigma}_{\rm c}}^{\bC\rarr\hat{\sA}\hat{\sB}}$.
The algorithm uses 
\begin{align}
    \zeta_{\tF} = \til{\cO}\Big(\f{d_\sA}{\delta^2\beta_\tF^3}\min\Big\{\f{1}{\rho_{\rm min}^2} + \f{1}{\sigma_{\rm min}^2}, \f{d_\sA^2}{\beta_\tF^4\delta^4}\Big\}\Big),
\end{align}
samples of $\rho^{\hat{\sA}}$ and $\sigma^{\hat{\sA}}$, where $\beta_\tF = \f{2}{\pi}\max\{s_{\rm min}, \delta/(8r)\}$.

In both cases, the quantum circuit for implementing the algorithm consists of $\cO\big(\zeta_\chi \log{d_\sA}\big)$ one- and two-qubit gates, and $\cO\big(\log{d_\sA}\big)$ qubits suffice at any one time. 
\end{theorem}




\begin{algorithm}[h]
\caption{Uhlmann transformation algorithm in the mixed sample access model \\ \parbox{\linewidth}{\centering $\mathtt{UhlmannMixedSample}(\rho^{\hat{\sA}}, \sigma^{\hat{\sA}}; \delta, \chi)$ (In Theorem~\ref{thm:Uhlmann mixed state sample})}}
\label{alg:Uhl mixed sample}
\SetKwInput{KwInput}{Input}
\SetKwInput{KwOutput}{Output}
\SetKwInput{KwParameters}{Parameters}
\SetKwComment{Comment}{$\triangleright$\ }{}
\SetCommentSty{textnormal}
\SetKwProg{Fn}{Subroutine}{}{end}
\SetKwFunction{UPS}{UhlmannPurifiedSample}

\SetAlgoNoEnd
\SetAlgoNoLine
\KwInput{Two quantum states $\rho^{\hat{\sA}}$ and $\sigma^{\hat{\sA}}$.}
\KwParameters{$\delta \in (0, 1)$ and $\chi \in \{\diamond, \tF\}$.}
\KwOutput{Quantum channel $\cJ^{\sB\sH\hat{\sA}\hat{\sB}}\circ\cP_{\til{\sigma}_{\rm c}}^{\bC\rarr\hat{\sA}\hat{\sB}}$.}
\SetAlgoLined

\If{$\chi = \diamond$}{
  Set $\delta_3 \gets (\delta/12)^2$ and $\beta \gets 2s_{\rm min}/\pi$. \\
}
\ElseIf{$\chi = \tF$}{
  Set $\delta_3 \gets \delta/16$ and $\beta \gets \f{1}{\pi}\max\{2s_{\rm min}, \delta_3/r\}$. \\
}
Set $u$ to be the minimum odd integer satisfying $u \geq \big\lceil\f{8e}{\beta}\log{(2/\delta_3)}\big\rceil$. \\
Set $\delta_1 \gets \delta/\big(2(4u+1)\big)$. \\
Set $\til{\rho}_{\rm c}^{\hat{\sA}\hat{\sB}} \gets \mathtt{CanonicalPurification}(\rho^{\hat{\sA}}; \delta_1)$ and $\til{\sigma}_{\rm c}^{\hat{\sA}\hat{\sB}} \gets \mathtt{CanonicalPurification}(\sigma^{\hat{\sA}}; \delta_1)$ (Algorithm~\ref{alg:canonical purification}). \\
Set $\cJ^{\sB\sH\hat{\sA}\hat{\sB}} \gets \mathtt{UhlmannPurifiedSample}(\til{\rho}_{\rm c}^{\hat{\sA}\hat{\sB}}, \til{\sigma}_{\rm c}^{\hat{\sA}\hat{\sB}}; \delta/2, \chi)$ (Algorithm~\ref{alg:Uhl purif sample}). \\
Set $\cP_{\til{\sigma}_{\rm c}}^{\bC\rarr\hat{\sA}\hat{\sB}} \gets \til{\sigma}_{\rm c}^{\hat{\sA}\hat{\sB}}$. \\
Return $\cJ^{\sB\sH\hat{\sA}\hat{\sB}}\circ\cP_{\til{\sigma}_{\rm c}}^{\bC\rarr\hat{\sA}\hat{\sB}}$. 
\end{algorithm}







In \Cref{alg:Uhl mixed sample}, $\mathtt{CanonicalPurification}$ is a quantum sample algorithm that prepares the states $\til{\rho}_{\rm c}^{\hat{\sA}\hat{\sB}}$ and $\til{\sigma}_{\rm c}^{\hat{\sA}\hat{\sB}}$, which approximate the canonical purified state $\ket{\rho_{\rm c}}^{\hat{\sA}\hat{\sB}}$ and $\ket{\sigma_{\rm c}}^{\hat{\sA}\hat{\sB}}$, respectively. We will introduce and explain this algorithm in detail in the next section, \Cref{sec:appendix B}. 

Note that, although $\mathtt{UhlmannPurifiedSample}$ is originally designed for pure state inputs, in this case, we allow general mixed states as inputs (see \Cref{alg:Uhl mixed sample}).
This requires a careful error analysis, which we will discuss in the proof of Theorem~\ref{thm:Uhlmann mixed state sample} provided in \Cref{sec:proof eval sample mixed}.






%===================================================================================================================================================================================






\subsection{Canonical purification algorithm}
\label{sec:appendix B}

We provide a quantum sample algorithm for the canonical purification $\mathtt{CanonicalPurification}$, which outputs an approximation of the canonical purified state $\ket{\omega_{\rm c}}^{\sA\sB}$ from multiple copies of $\omega^\sA$. 
The details of $\mathtt{CanonicalPurification}$ are provided in \Cref{alg:canonical purification}, where $\sF$ and $\sG$ are five- and six-qubit systems, respectively.
A quantum sample algorithm $\mathtt{BlockEncSqrtState}$ is described in \Cref{prop:block enc sqrt state} below.
Regarding the performance of $\mathtt{CanonicalPurification}$, the following proposition holds.


\begin{theorem}[Quantum algorithm for the canonical purification]
\label{prop:canonical purification}
Let $\delta \in (0, 1)$, and let $\omega^\sA$ be a state whose minimum non-zero eigenvalue is $\omega_{\rm min}$. Then, a quantum sample algorithm $\mathtt{CannonicalPurification}(\omega^\sA; \delta)$ given by \Cref{alg:canonical purification} outputs a state $\til{\omega}_{\rm c}^{\sA\sB}$ such that
\begin{align}
    \f{1}{2}\|\til{\omega}_{\rm c}^{\sA\sB} - \ketbra{\omega_{\rm c}}{\omega_{\rm c}}^{\sA\sB}\|_1 \leq \delta,
\end{align}
using $f = \til{\cO}\Big(\f{d_\sA}{\delta}\min\Big\{\f{1}{\omega_{\rm min}^2}, \f{d_\sA^2}{\delta^4}\Big\}\Big)$ samples of $\omega^\sA$.
The quantum circuit for the algorithm is constructed by $\cO(f\log d_\sA)$ one- and two-qubit gates, and at any one time, $\cO(\log{d_\sA})$ qubits suffice.
\end{theorem}



\begin{algorithm}[h]
\caption{Canonical purification algorithm \\ \parbox{\linewidth}{\centering $\mathtt{CanonicalPurification}(\omega^\sA; \delta)$ (In Theorem~\ref{prop:canonical purification})}}
\label{alg:canonical purification}
\SetKwInput{KwInput}{Input}
\SetKwInput{KwOutput}{Output}
\SetKwInput{KwParameters}{Parameters}
\SetKwComment{Comment}{$\triangleright$\ }{}
\SetCommentSty{textnormal}

\SetAlgoNoEnd
\SetAlgoNoLine
\KwInput{Quantum state $\omega^\sA$.}
\KwParameters{$\delta \in (0, 1)$.}
\KwOutput{Quantum state $\til{\omega}_{\rm c}^{\sA\sB}$.}
\SetAlgoLined


Set $\delta_1 \gets (\delta/6)^2$ and $\beta_\diamond \gets 1/(2\sqrt{2d_\sA})$. \\
Set $u$ to be the minimum odd integer satisfying $u \geq \big\lceil\f{8e}{\beta_\diamond}\log{(2/\delta_1)}\big\rceil$. \\


Set $\cF^{\sA\sF} \gets \mathtt{BlockEncSqrtState}(\omega^\sA; \delta_1/(2u))$ and $\cF_{\rm inv}^{\sA\sF} \gets \mathtt{BlockEncSqrtState}^\dag(\omega^\sA; \delta_1/(2u))$ (Proposition~\ref{prop:block enc sqrt state}). \\
Set $\cG^{\sA\sB\sF} \gets \cF^{\sA\sF}\circ\cU_{\Phi}^{\sA\sB}$ and $\cG_{\rm inv}^{\sA\sB\sF} \gets (\cU_{\Phi}^{\sA\sB})^\dag\circ\cF_{\rm inv}^{\sA\sF}$, where $U_{\Phi}^{\sA\sB}$ is a unitary preparing the maximally entangled state: $U_{\Phi}^{\sA\sB}\ket{0}^{\sA\sB} = \ket{\Phi}^{\sA\sB}$. \\
Set $\cL^{\sA\sB\sG} \gets \mathtt{QSVTSIGN}(\cG^{\sA\sB\sF}, \cG_{\rm inv}^{\sA\sB\sF}; \delta_1, \beta_\diamond)$ (Proposition~\ref{prop:FPAA general}). \\
Set $\til{\omega}_{\rm c}^{\sA\sB} \gets \tr_\sG\circ\cL^{\sA\sB\sG}\circ\cP_{\ket{0}}^{\bC\rarr\sG}(\ketbra{0}{0}^{\sA\sB})$. \\
Return $\til{\omega}_{\rm c}^{\sA\sB}$.
\end{algorithm}



Our approach begins by approximately constructing a unitary, which is a block-encoding of $\sqrt{\omega^\sA}$, using $\mathtt{BlockEncSqrtState}$.
We then apply the unitary to the state $\ket{0}^{\sM}\ket{\Phi}^{\sA\sB}$ and amplify the coefficient to unity by the QSVT with the sign function.



\begin{proposition}
\label{prop:block enc sqrt state}
Let $\delta \in (0, 1)$, and let $\omega$ be a state in a $d$-dimensional Hilbert space, whose minimum non-zero eigenvalue is $\omega_{\rm min}$. 
Then, there is a quantum sample algorithm $\mathtt{BlockEncSqrtState}(\omega; \delta)$ that outputs a channel $\cF$ such that $\f{1}{2}\|\cF - \cU\|_\diamond \leq \delta$, using $h = \til{\cO}\big(\f{1}{\delta}\min\big\{1/\omega_{\rm min}^2, 1/\delta^4\big\}\big)$ samples of $\omega$,
where $U$ is a $(1, 5, 0)$-block-encoding unitary of $\sqrt{\omega}/\big(2\sqrt{2}\big)$.
The quantum circuit for the algorithm consists of $\cO(h\log{d})$ one- and two-qubit gates, and $\cO(\log{d})$ qubits suffice at any one time.

Moreover, there is a quantum sample algorithm $\mathtt{BlockEncSqrtState}^\dag(\omega; \delta)$ that outputs a channel $\cF_{\rm inv}$ such that $\f{1}{2}\|\cF_{\rm inv} - \cU^\dag\|_\diamond \leq \delta$, using the same number of samples, gates, and qubits.
\end{proposition}




We first prove \Cref{prop:block enc sqrt state}.    
At a high level, this is achieved by the following two steps:
\begin{align}
\label{eq:flow of trans purif}
\omega^\sA \overset{1. \, \text{Theorem~\ref{thm:BEquantum state form sample}}}{\longrightarrow}
\begin{pmatrix}
    \omega^\sA & \hspace*{1.5mm} * \hspace*{3mm} \\
    * & \hspace{1.5mm} * \hspace*{3mm}
\end{pmatrix}
\overset{2. \, \text{Proposition~\ref{prop:QSVT sqrt}}}{\longrightarrow}
\begin{pmatrix}
    \sqrt{\omega^\sA} & \hspace*{1.5mm} * \hspace*{3mm} \\
    * & \hspace{1.5mm} * \hspace*{3mm}
\end{pmatrix}.
\end{align}




In step~1, we use quantum algorithms that block-encode the density matrix of a quantum state \cite{gilyen2022improvedfidelity, wang2024fastQAtracedist, wang2023SampletoQuary, wang2024TimeEfficientEntEstSamp}, which are obtained by combining the density matrix exponentiation and the QSVT. 

\begin{theorem}[Block-encoding of a quantum state from samples~{\cite[Lemma 2.21]{wang2023SampletoQuary}}]
\label{thm:BEquantum state form sample}
Let $\delta \in (0, 1)$, and let $\omega$ be a state in a $d$-dimensional Hilbert
space. 
Then, there are quantum sample algorithms that output channels $\cE$ and $\cE_{\rm inv}$, respectively, which satisfy $\f{1}{2}\|\cE - \cU\|_\diamond \leq \delta$ and $\f{1}{2}\|\cE_{\rm inv} - \cU^\dag\|_\diamond \leq \delta$, where $U$ is a $(1, 4, 0)$-block-encoding unitary of $\omega/2$.
Both algorithms use $y$ samples of $\omega$, where $y = \cO\big(\f{1}{\delta}\big(\log{\big(\f{1}{\delta}\big)}\big)^2\big)$.
The quantum circuits for these algorithms consist of $\cO(y\log{d})$ one- and two-qubit gates, and $\cO(\log{d})$ qubits suffice at any one time.
\end{theorem}


In step~2, we use the QSVT to approximate the square root function. 
Specifically, we take the target function for a polynomial approximation to be $\sqrt{x}/2$. This is a special case of the QSVT for positive power functions; see, e.g., Refs.~\cite{Gilyn2019thesis, gilyen2022improvedfidelity, wang2024NewQuantEnt} for details.



\begin{proposition}[QSVT with the square root function~{\cite[Lemma 17 and Corollary 18]{gilyen2022improvedfidelity}},~{\cite[Lemma II.6 and Lemma II.13]{wang2024NewQuantEnt}}]
\label{prop:QSVT sqrt}
Let $\delta \in (0, 1/2)$, and let $U$ be a $(1, a, 0)$-block-encoding unitary of a positive semidefinite matrix $M$ whose minimum non-zero eigenvalue is $m_{\rm min}$. Then, there are quantum algorithms that output unitaries $\til{U}$ and $\til{U}^\dag$, respectively, where $\til{U}$ is a $(1, a+1, 0)$-block-encoding of an even/odd polynomial $R_{\rm sqrt}(M)$.
The polynomial $R_{\rm sqrt}$ satisfies $\big\|R_{\rm sqrt}(M) - \sqrt{M}/2\big\|_\infty \leq \delta$, and $|R_{\rm sqrt}(x)| \leq 1$ for $x \in [-1, 1]$. 
Both algorithms use $U$ and $U^\dag$ a total of $l = \cO\big(\min\{1/m_{\rm min}, 1/\delta^2\}\log{(1/\delta)}\big)$ times.
\end{proposition}

This proposition follows from the existence of polynomials $P_{\rm sqrt}$ and $Q_{\rm sqrt}$, each of which is either even or odd, with degrees $\cO\big(\f{1}{\alpha}\log{(1/\delta)}\big)$ and $\cO\big(\f{1}{\delta^2}\log{(1/\delta)}\big)$, respectively, where $\alpha \in (0, 1/2)$.
These polynomials satisfy that 
\begin{align}
    &\big|P_{\rm sqrt}(x)\big|, \big|Q_{\rm sqrt}(x)\big| \leq 1 
    \ \ \textrm{for} \ \ x \in [-1, 1], \\
    &\Big|P_{\rm sqrt}(x) - \sqrt{x}/2\Big| \leq \delta
    \ \ \textrm{for} \ \ x \in \big[\alpha/2, 1\big], \\
    &\Big|Q_{\rm sqrt}(x) - \sqrt{x}/2\Big| \leq \delta
    \ \ \textrm{for} \ \ x \in [-1, 1].
\end{align}
For our purpose, it suffices to set $\alpha$ so that $\alpha/2 = m_{\rm min}$.



\begin{proof}[Proof of \Cref{prop:block enc sqrt state}]


Let $\delta_1 \in (0, 1)$, and $\sE$ be a four-qubit system.
From Theorem~\ref{thm:BEquantum state form sample}, we can obtain quantum channels $\cE^{\sA\sE}$ and $\cE_{\rm inv}^{\sA\sE}$, which satisfy that
\begin{align}
    \f{1}{2}\big\|\cE^{\sA\sE} - \cU_1^{\sA\sE}\big\|_\diamond \leq \delta_1, \ \ \text{and} \ \ \ 
    \f{1}{2}\big\|\cE_{\rm inv}^{\sA\sE} - (\cU_1^{\sA\sE})^\dag\big\|_\diamond \leq \delta_1,
\end{align}
where $U_1^{\sA\sE}$ is a block- encoding unitary of $\omega/2$, using $y = \cO\Big(\big(\log{(1/\delta_1)}\big)^2/\delta_1\Big)$ samples of $\omega^\sA$.


Let $\delta_2 \in (0, 1/2)$, $\omega_{\rm min}$ be the minimum non-zero eigenvalue of $\omega^\sA$, and $\sF$ be a five-qubit system. 
We evaluate the accumulated error that arises from using quantum channels $\cE^{\sA\sE}$ and $\cE_{\rm inv}^{\sA\sE}$ instead of a unitary $U_1^{\sA\sE}$ and $(U_1^{\sA\sE})^\dag$ in the QSVT with the square root function.
From \Cref{prop:QSVT sqrt}, by using $\cE^{\sA\sE}$ and $\cE_{\rm inv}^{\sA\sE}$ for a total of $l = \cO\big(\min\{1/\omega_{\rm min}, 1/\delta_2^2\}\log{(1/\delta_2)}\big)$ times, we can obtain quantum channels $\cF^{\sA\sF}$ and $\cF_{\rm inv}^{\sA\sF}$ such that 
\begin{align}
\label{inteq:93}
    \f{1}{2}\big\|\cF^{\sA\sF} - \cU_2^{\sA\sF}\big\|_\diamond \leq l\delta_1, \ \  \text{and} \ \ \
    \f{1}{2}\big\|\cF_{\rm inv}^{\sA\sF} - (\cU_2^{\sA\sF})^\dag\big\|_\diamond \leq l\delta_1,
\end{align}
where we used the subadditivity of the diamond norm to bound the overall error.
The unitary $U_2^{\sA\sF}$ is a block-encoding unitary of a polynomial $R_{\rm sqrt}(\omega^\sA/2)$, which satisfies that 
\begin{align}
    \big\|R_{\rm sqrt}(\omega^\sA/2) - \sqrt{\omega^\sA}/\big(2\sqrt{2}\big)\big\|_\infty \leq \delta_2.
\end{align}


To evaluate the distance, in terms of the diamond norm, between $\cU_2^{\sA\sF}$ and $\cU_{\rm ideal}^{\sA\sF}$, where $U_{\rm ideal}^{\sA\sF}$ is an exact block-encoding unitary of $\sqrt{\omega^\sA}/\big(2\sqrt{2}\big)$, we use the following lemma~\cite{Gilyn2019thesis, gilyen2022improvedfidelity, wang2023SampletoQuary}.

\begin{lemma}[Robustness of block-encoding unitaries~{\cite[Corollary 12]{gilyen2022improvedfidelity}}]
\label{prevthm:robustness BE unitary}
Suppose that $U$ is a $d \times d$ unitary that is block-encoding of a $m' \times m$ matrix $A$ and $d \geq 4(m'+m)$. Then, for any $m' \times m$ matrix $\til{A}$ satisfying $\|A-\til{A}\|_\infty + \Big\|\f{A+\til{A}}{2}\Big\|_\infty^2 \leq 1$, there exists a $d \times d$ unitary $\til{U}$ that is block-encoding of $\til{A}$ satisfying
\begin{align}
    \big\|U - \til{U}\big\|_\infty &\leq \sqrt{\f{2}{1-\big\|\f{A+\til{A}}{2}\big\|_\infty^2}}\big\|A-\til{A}\big\|_\infty.
\end{align}  
\end{lemma}


Observe that, for $\delta_2 \in (0, 1/2)$,
\begin{align}
    \big\|R_{\rm sqrt}(\omega^\sA/2) - \sqrt{\omega^\sA}/\big(2\sqrt{2}\big)\big\|_\infty + \Big\|\f{1}{2}\Big(R_{\rm sqrt}(\omega^\sA/2) + \sqrt{\omega^\sA}/\big(2\sqrt{2}\big)\Big)\Big\|_\infty^2 
    &\leq \delta_2 + \f{1}{4}\big(2\big\|\sqrt{\omega^\sA}/\big(2\sqrt{2}\big)\big\|_\infty + \delta_2\big)^2 \\
    &\leq \delta_2 + \f{1}{4}\big(1/\sqrt{2} + \delta_2\big)^2 \\
    &< 1/2 + \f{1}{4}\big(1/\sqrt{2} + 1/2\big)^2 \\
    &< 1.
\end{align}
Thus, applying \Cref{prevthm:robustness BE unitary}, it holds that 
\begin{equation}
\label{inteq:94}
    \f{1}{2}\big\|\cU_2^{\sA\sF} - \cU_{\rm ideal}^{\sA\sF}\big\|_\diamond \leq 2 \delta_2,
\end{equation}
where we used $1-\Big\|\f{1}{2}\Big(R_{\rm sqrt}(\omega^\sA/2) + \sqrt{\omega^\sA}/\big(2\sqrt{2}\big)\Big)\Big\|_\infty^2 \geq 1/2$ and the relation between the diamond norm and the operator norm in Eq.~\eqref{eq:rel of operator and diamond norm}.
From Eqs.~\eqref{inteq:93} and~\eqref{inteq:94}, we have that $\f{1}{2}\big\|\cF^{\sA\sF} - \cU_{\rm ideal}^{\sA\sF}\big\|_\diamond \leq l\delta_1 + 2\delta_2$.
Similarly, we have $\f{1}{2}\big\|\cF_{\rm inv}^{\sA\sF} - (\cU_{\rm ideal}^{\sA\sF})^\dag\big\|_\diamond \leq l\delta_1 + 2\delta_2$.
Rescaling the parameters as $\delta_1 = \delta/(2l)$, $\delta_2 = \delta/4$, the overall errors are bounded by $\delta$.

The number of samples of $\omega^\sA$ required so far is evaluated as
\begin{align}
    h 
    &= yl \\ 
    &= \cO\Big(\f{1}{\delta_1}\Big(\log{\Big(\f{1}{\delta_1}\Big)}\Big)^2\Big) \cO\Big(\min\Big\{\f{1}{\omega_{\rm min}}, \f{1}{\delta_2^2}\Big\}\log{\Big(\f{1}{\delta_2}\Big)}\Big) \\
    &=\cO\bigg(\f{1}{\delta}\min\Big\{\f{1}{\omega_{\rm min}^2}, \f{1}{\delta^4}\Big\} \Big(\log{\Big(\f{1}{\delta}\Big)}\Big)^2 \Big(\log{\Big(\f{1}{\delta}\min\Big\{\f{1}{\omega_{\rm min}}, \f{1}{\delta^2}\Big\}\log\Big(\f{1}{\delta}\Big)}\Big)\Big)^2\bigg) \\
    &=\til{\cO}\Big(\f{1}{\delta}\min\Big\{\f{1}{\omega_{\rm min}^2}, \f{1}{\delta^4}\Big\}\Big).
\end{align}
The number of one- and two-qubit gates in the quantum circuit of this algorithm is given by $\cO\big(h\log{d_\sA}\big)$.
Since this algorithm is conducted sequentially, $\cO(\log{d_\sA})$ qubits suffice at any one time.

\end{proof}



We now provide a proof of Theorem~\ref{prop:canonical purification}.


\begin{proof}[Proof of Theorem~\ref{prop:canonical purification}]
    

Let $U_{\rm ideal}^{\sA\sF}$ be a (1, 5, 0)-block-encoding unitary of $\sqrt{\omega^\sA}/\big(2\sqrt{2}\big)$, and let $U_{\Phi}^{\sA\sB}$ be a unitary that prepares the maximally entangled state $\ket{\Phi}^{\sA\sB}$: $U_{\Phi}^{\sA\sB}\ket{0}^{\sA\sB} = \ket{\Phi}^{\sA\sB}$.
We can observe that
\begin{align}
    \bra{0}^{\sF}U_{\rm ideal}^{\sA\sF}U_{\Phi}^{\sA\sB}\ket{0}^{\sF}\ket{0}^{\sA\sB} 
    &= \f{\sqrt{\omega^\sA}}{2\sqrt{2}}\ket{\Phi}^{\sA\sB} \\
    &= \f{1}{2
    \sqrt{2d_\sA}}\ket{\omega_{\rm c}}^{\sA\sB}.
\end{align}
This implies that the unitary $U_{\rm ideal}^{\sA\sF}U_{\Phi}^{\sA\sB}$ is a block encoding of $\f{1}{2\sqrt{2d_\sA}}\ket{\omega_{\rm c}}^{\sA\sB}$, i.e., 
\begin{align}
\label{inteq:95}
    U_{\rm ideal}^{\sA\sF}U_{\Phi}^{\sA\sB}
    = \hspace{1mm}
\begin{blockarray}{ccc}
& \bra{0}^\sF\bra{0}^{\sA\sB}&  & \vspace{1mm}\\
\begin{block}{c(cc)}
  \ket{0}^\sF \hspace*{1mm} & \f{1}{2\sqrt{2d_\sA}}\ket{\omega_{\rm c}}^{\sA\sB} & \hspace{0mm} * \hspace{3mm}\\
   \hspace*{1mm} & * & \hspace{0mm} * \hspace{3mm}\\
\end{block}
\end{blockarray}\hspace{2.5mm}.
\end{align}
By the QSVT with the sign function in \Cref{prop:FPAA general}, we amplify the coefficient $1/(2\sqrt{2d_\sA})$ to unity.
To this end, we set $\beta$ to $\beta_\diamond = 1/(2\sqrt{2d_\sA})$.
Note that, on the right-hand side of Eq.~\eqref{inteq:95}, the states $\ket{0}^\sF$ and $\ket{0}^\sF\ket{0}^{\sA\sB}$, which specify the encoded block, do not have the same dimensions. Even in this case, the QSVT can still be performed without any issue~\cite{gilyen2019qsvt, Gilyn2019thesis, martyn2021grand}.


We denote by $\cF^{\sA\sF}$ and $\cF_{\rm inv}^{\sA\sF}$ the quantum channels given by $\cF^{\sA\sF} = \mathtt{BlockEncSqrtState}(\omega^\sA; \delta_1)$ and $\cF_{\rm inv}^{\sA\sF} = \mathtt{BlockEncSqrtState}^\dag(\omega^\sA; \delta_1)$ for $\delta_1 \in (0, 1)$, respectively.
Proposition~\ref{prop:block enc sqrt state} ensures that
\begin{align}
    \f{1}{2}\big\|\cF^{\sA\sF} - \cU_{\rm ideal}^{\sA\sF}\big\|_\diamond \leq \delta_1, \ \ \text{and} \ \ \ \f{1}{2}\big\|\cF_{\rm inv}^{\sA\sF} - (\cU_{\rm ideal}^{\sA\sF})^\dag\big\|_\diamond \leq \delta_1,
\end{align}
when we use $h = \til{\cO}\big(\f{1}{\delta_1}\min\big\{\f{1}{\omega_{\rm min}^2}, \f{1}{\delta_1^4}\big\}\big)$ samples of $\omega^\sA$.

We evaluate the error originated by using the quantum channel $\cG^{\sA\sB\sF} = \cF^{\sA\sF}\circ\cU_{\Phi}^{\sA\sB}$ and $\cG_{\rm inv}^{\sA\sB\sF} = (\cU_{\Phi}^{\sA\sB})^\dag \circ \cF_{\rm inv}^{\sA\sF}$, instead of $U_{\rm ideal}^{\sA\sF}U_{\Phi}^{\sA\sB}$ and $(U_{\rm ideal}^{\sA\sF}U_{\Phi}^{\sA\sB})^\dag$, in the input of $\mathtt{QSVTSIGN}$.
From \Cref{lem:FPAA diamond}, the subadditivity of the diamond norm, and the triangle inequality, we see that a quantum channel $\cL^{\sA\sB\sG} = \mathtt{QSVTSIGN}(\cG^{\sA\sB\sF}, \cG_{\rm inv}^{\sA\sB\sF}; \delta_2, \beta_\diamond)$ for $\delta_2 \in (0, 1/2)$ satisfies
\begin{align}
    &\f{1}{2}\big\|\cL^{\sA\sB\sG} \circ \cP_{\ket{0}}^{\bC\rarr\sG} - \cW_{\rm ideal}^{\sA\sB\sG} \circ\cP_{\ket{0}}^{\bC\rarr\sG}\|_\diamond \leq 3\sqrt{\delta_2} + u\delta_1,    
\end{align}
where $W_{\rm ideal}^{\sA\sB\sG}$ is an exact block-encoding unitary of $\ket{\omega_{\rm c}}^{\sA\sB}$.
Here, we use $\cG^{\sA\sB\sF}$ and $\cG_{\rm inv}^{\sA\sB\sF}$ $u$ times in total, where $u$ is the minimum odd integer such that $u \geq \big\lceil\f{8e}{\beta_\diamond}\log(2/\delta_2)\big\rceil = \big\lceil16e\sqrt{2d_\sA}\log(2/\delta_2)\big\rceil$.



Noting that $\tr_\sG\circ\cW_{\rm ideal}^{\sA\sB\sG} \circ\cP_{\ket{0}}^{\bC\rarr\sG}(\ketbra{0}{0}^{\sA\sB}) = \ketbra{\omega_{\rm c}}{\omega_{\rm c}}^{\sA\sB}$, we obtain a quantum state $\til{\omega}_{\rm c}^{\sA\sB} = \tr_\sG\circ\cL^{\sA\sB\sG}\circ\cP_{\ket{0}}^{\bC\rarr\sG}(\ketbra{0}{0}^{\sA\sB})$ such that
\begin{align}
    \f{1}{2}\big\|\til{\omega}_{\rm c}^{\sA\sB} - \ketbra{\omega_{\rm c}}{\omega_{\rm c}}^{\sA\sB}\big\|_1 \leq 3\sqrt{\delta_2} + u\delta_1,
\end{align}
where we used the definition of the diamond norm and the contraction property of the trace norm under the partial trace.
Rescaling the parameters as $\delta_1 = \delta/(2u)$ and $\delta_2 = (\delta/6)^2$, the overall error can be bounded by $\delta$.


The number of samples of $\omega^\sA$ required by the entire algorithm is evaluated as
\begin{align}
    f 
    &= hu \\ 
    &= \til{\cO}\Big(\f{1}{\delta_1}\min\Big\{\f{1}{\omega_{\rm min}^2}, \f{1}{\delta_1^4}\Big\}\Big)\cO\Big(\sqrt{d_\sA}\log{\Big(\f{1}{\delta_2}\Big)}\Big) \\
    &=\til{\cO}\Big(\f{d_\sA}{\delta}\min\Big\{\f{1}{\omega_{\rm min}^2}, \f{d_\sA^2}{\delta^4}\Big\}\Big).
\end{align}
The number of one- and two-qubit gates in the quantum circuit of this algorithm is evaluated as $\cO\big(f\log{d_\sA}\big)$.
Since this algorithm is conducted sequentially, $\cO(\log{d_\sA})$ qubits suffice at any one time.




\end{proof}






%===========================================================================================================================================================================================================================================================






\subsection{Proof of Theorem~\ref{thm:Uhlmann mixed state sample}}
\label{sec:proof eval sample mixed}



Using $\mathtt{CanonicalPurification}$ with $\delta_1 \in (0, 1)$, we obtain states $\til{\rho}_{\rm c}^{\hat{\sA}\hat{\sB}}$ and $\til{\sigma}_{\rm c}^{\hat{\sA}\hat{\sB}}$ such that
\begin{align}
\label{inteq:41}
    &\f{1}{2}\big\|\til{\rho}_{\rm c}^{\hat{\sA}\hat{\sB}} - \ketbra{\rho_{\rm c}}{\rho_{\rm c}}^{\hat{\sA}\hat{\sB}}\big\|_1 \leq \delta_1, \ \ \text{and} \ \ \ \f{1}{2}\big\|\til{\sigma}_{\rm c}^{\hat{\sA}\hat{\sB}} - \ketbra{\sigma_{\rm c}}{\sigma_{\rm c}}^{\hat{\sA}\hat{\sB}}\big\|_1 \leq \delta_1,
\end{align}
from $f$ samples of $\rho^{\hat{\sA}}$ and $\sigma^{\hat{\sA}}$, where 
\begin{align}
\label{inteq:43}
    f = \til{\cO}\Big(\f{d_\sA}{\delta_1}\min\Big\{\f{1}{\rho_{\rm min}^2} + \f{1}{\sigma_{\rm min}^2}, \f{d_\sA^2}{\delta_1^4}\Big\}\Big),
\end{align}
and $\rho_{\rm min}$ and $\sigma_{\rm min}$ are minimum non-zero eigenvalues of $\rho^\sA$ and $\sigma^\sA$, respectively. Here, we used that 
\begin{align}
    \min\{1/\rho_{\rm min}^2, d_\sA^2/\delta_1^4\} + \min\{1/\sigma_{\rm min}^2, d_\sA^2/\delta_1^4\}  
    &\leq \min\{1/\rho_{\rm min}^2 + 1/\sigma_{\rm min}^2, 2d_\sA^2/\delta_1^4\} \\
    &= \cO(\min\{1/\rho_{\rm min}^2 + 1/\sigma_{\rm min}^2, d_\sA^2/\delta_1^4\}).
\end{align}
We need to evaluate how the error $\delta_1$ in Eqs.~\eqref{inteq:41} affect the result, when we use $\til{\rho}_{\rm c}^{\hat{\sA}\hat{\sB}}$ and $\til{\sigma}_{\rm c}^{\hat{\sA}\hat{\sB}}$ as inputs to $\mathtt{UhlmannPurifiedSample}$, instead of pure states. 


When we run the circuit in \Cref{fig:upsilon circuit}, replacing $\ket{\rho}^{\sA_1\sB_1}$ and $\ket{\sigma}^{\sA_2\sB_2}$ with $\til{\rho}_{\rm c}^{\sA_1\sB_1}$ and $\til{\sigma}_{\rm c}^{\sA_2\sB_2}$, respectively, the output state is given by 
\begin{align}
    &\til{\Upsilon}^{\sC_2\sC_3\sA_1\sB_1\sB_2}
    =\f{1}{2}\big(\ketbra{0}{0}^{\sC_2} \otimes \til{\xi}^{\sC_3\sA_1\sB_1\sB_2} + \ketbra{1}{1}^{\sC_2} \otimes Z^{\sC_3}\til{\xi}^{\sC_3\sA_1\sB_1\sB_2}Z^{\sC_3}\big),
\end{align}
where $\til{\xi}^{\sC_3\sA_1\sB_1\sB_2}$ is given by
\begin{align}
    \til{\xi}^{\sC_3\sA_1\sB_1\sB_2} 
    = \hspace{1mm} \f{1}{2}
\begin{blockarray}{ccc}
& \bra{0}^{\sC_3}&  \bra{1}^{\sC_3} \vspace{1mm}\\
\begin{block}{c(cc)}
  \ket{0}^{\sC_3}\hspace*{1mm} & \til{\sigma}_{\rm c}^{\sA_1\sB_1}\otimes \til{\rho}_{\rm c}^{\sB_2} & \big(\til{L}^{\sA_1\sB_1\sB_2}\big)^\dag \\
  \ket{1}^{\sC_3}\hspace*{1mm} & \til{L}^{\sA_1\sB_1\sB_2} & \til{\rho}_{\rm c}^{\sA_1\sB_1}\otimes \til{\sigma}_{\rm c}^{\sB_2} \\
\end{block}
\end{blockarray}\hspace{2.5mm},
\end{align}
and $\til{L}^{\sA_1\sB_1\sB_2} = \tr_{\sA_2}\big[(\til{\rho}_{\rm c}^{\sA_1\sB_1}\otimes \til{\sigma}_{\rm c}^{\sA_2\sB_2})F^{\sA_1\sA_2\sB_1\sB_2}\big]$.
From \Cref{prevthm:DME variant}, for $\delta_2 \in (0, 1)$, we obtain a quantum channel $\til{\cF}^{\sC\hat{\sA}\hat{\sB}\sB} = \mathtt{DMESUB}(\til{\Upsilon}^{\sC_2\sC_3\sA_1\sB_1\sB_2}; \delta_2, t=2)$, which satisfies that
\begin{align}
\label{inteq:46}
    \f{1}{2}\big\|\til{\cF}^{\sC\hat{\sA}\hat{\sB}\sB} - \til{\cU}_2^{\sC\hat{\sA}\hat{\sB}\sB}\big\|_\diamond \leq \delta_2,
\end{align}
where $\til{U}_2^{\sC\hat{\sA}\hat{\sB}\sB} = e^{i\til{K}^{\sC\hat{\sA}\hat{\sB}\sB}}$
and $\til{K}^{\sC\hat{\sA}\hat{\sB}\sB} = \ketbra{1}{0}^\sC \otimes \til{L}^{\hat{\sA}\hat{\sB}\sB} + \ketbra{0}{1}^\sC \otimes (\til{L}^{\hat{\sA}\hat{\sB}\sB})^\dag$, using $m = \cO(1/\delta_2)$ samples of $\til{\rho}_{\rm c}^{\hat{\sA}\hat{\sB}}$ and $\til{\sigma}_{\rm c}^{\hat{\sA}\hat{\sB}}$.



We evaluate the distance between $\til{U}_2^{\sC\hat{\sA}\hat{\sB}\sB}$ and the desired unitary $U_2^{\sC\hat{\sA}\hat{\sB}\sB} = e^{iK^{\sC\hat{\sA}\hat{\sB}\sB}}$,
where $K^{\sC\hat{\sA}\hat{\sB}\sB} = \ketbra{1}{0}^\sC \otimes L^{\hat{\sA}\hat{\sB}\sB} + \ketbra{0}{1}^\sC \otimes (L^{\hat{\sA}\hat{\sB}\sB})^\dag$,
and
\begin{align}
    L^{\hat{\sA}\hat{\sB}\sB} 
    &= \ketbra{\rho_{\rm c}}{\sigma_{\rm c}}^{\hat{\sA}\hat{\sB}} \otimes \tr_{\sA'}\big[\ketbra{\sigma_{\rm c}}{\rho_{\rm c}}^{\sA'\sB}\big] \\
    &= \tr_{\sA'}\big[(\ketbra{\rho_{\rm c}}{\rho_{\rm c}}^{\hat{\sA}\hat{\sB}} \otimes \ketbra{\sigma_{\rm c}}{\sigma_{\rm c}}^{\sA'\sB})F^{\hat{\sA}\sA'\hat{\sB}\sB}\big].
\end{align}
To this end, we use the following lemma shown in Ref.~\cite{chakraborty2019powerblock}. We provide a proof in \Cref{sec:appendix error not accumulate} for completeness.




\begin{restatable}[Robustness of the exponential function of an Hermitian matrix~{\cite[Lemma 50 in its full version]{chakraborty2019powerblock}}]{lemma}{lemrobstexp}
\label{lem:state error not accumulate}
Let $A$ and $B$ be Hermitian matrices.
Then, for any $t \in \bR$, it holds that
\begin{align}
    &\big\|e^{itA} - e^{itB}\big\|_\infty \leq |t|\|A - B\|_\infty.
\end{align}
\end{restatable}



From \Cref{lem:state error not accumulate} and the fact that $\|\til{K}^{\sC\hat{\sA}\hat{\sB}\sB} - K^{\sC\hat{\sA}\hat{\sB}\sB}\|_\infty = \|\til{L}^{\sC\hat{\sA}\hat{\sB}\sB} - L^{\sC\hat{\sA}\hat{\sB}\sB}\|_\infty$, we observe
\begin{align}
    \big\|\til{U}_2^{\sC\hat{\sA}\hat{\sB}\sB} - U_2^{\sC\hat{\sA}\hat{\sB}\sB}\big\|_\infty
    &\leq \big\|\til{K}^{\sC\hat{\sA}\hat{\sB}\sB} - K^{\sC\hat{\sA}\hat{\sB}\sB}\big\|_\infty \\
    \label{inteq:147}
    &= \|\til{L}^{\hat{\sA}\hat{\sB}\sB} - L^{\hat{\sA}\hat{\sB}\sB}\|_\infty.
\end{align}
We further compute as follows:
\begin{align}
    \big\|\til{L}^{\hat{\sA}\hat{\sB}\sB} -L^{\hat{\sA}\hat{\sB}\sB}\big\|_\infty 
    &\leq \big\|\til{L}^{\hat{\sA}\hat{\sB}\sB} - L^{\hat{\sA}\hat{\sB}\sB}\big\|_1 \\
    &\leq \big\|\til{\rho}_{\rm c}^{\hat{\sA}\hat{\sB}} \otimes \til{\sigma}_{\rm c}^{\sA'\sB} - \ketbra{\rho_{\rm c}}{\rho_{\rm c}}^{\hat{\sA}\hat{\sB}} \otimes \ketbra{\sigma_{\rm c}}{\sigma_{\rm c}}^{\sA'\sB}\big\|_1 \\
    &\leq \big\|\til{\rho}_{\rm c}^{\hat{\sA}\hat{\sB}} - \ketbra{\rho_{\rm c}}{\rho_{\rm c}}^{\hat{\sA}\hat{\sB}}\big\|_1 + \big\|\til{\sigma}_{\rm c}^{\sA'\sB} - \ketbra{\sigma_{\rm c}}{\sigma_{\rm c}}^{\sA'\sB}\big\|_1 \\
    \label{inteq:148}
    &\leq 4\delta_1,
\end{align}
where we used the contraction property of trace norm Eq.~\eqref{eq:contraction trace norm} and its isometric invariance in the second inequality, and used Eqs.~\eqref{inteq:41} in the last inequality.
Thus, we have 
\begin{align}
    \f{1}{2}\big\|\til{\cU}_2^{\sC\hat{\sA}\hat{\sB}\sB} - \cU_2^{\sC\hat{\sA}\hat{\sB}\sB}\big\|_\diamond
    &\leq \big\|\til{U}_2^{\sC\hat{\sA}\hat{\sB}\sB} - U_2^{\sC\hat{\sA}\hat{\sB}\sB}\big\|_\infty \\
    &\leq\|\til{L}^{\hat{\sA}\hat{\sB}\sB} - L^{\hat{\sA}\hat{\sB}\sB}\|_\infty \\
    \label{inteq:62}
    &\leq 4\delta_1,
\end{align}
where we used the relation between the diamond norm and operator norm in Eq.~\eqref{eq:rel of operator and diamond norm} and Eqs.~\eqref{inteq:147} and~\eqref{inteq:148}.


Using Eqs.~\eqref{inteq:46} and~\eqref{inteq:62}, we observe that
\begin{align}
\label{inteq:98}
    \f{1}{2}\big\|\til{\cF}^{\sC\hat{\sA}\hat{\sB}\sB} - \cU_2^{\sC\hat{\sA}\hat{\sB}\sB}\big\|_\diamond
    &\leq \f{1}{2}\big\|\til{\cF}^{\sC\hat{\sA}\hat{\sB}\sB} - \til{\cU}_2^{\sC\hat{\sA}\hat{\sB}\sB}\big\|_\diamond 
    + \f{1}{2}\big\|\til{\cU}_2^{\sC\hat{\sA}\hat{\sB}\sB} - \cU_2^{\sC\hat{\sA}\hat{\sB}\sB}\big\|_\diamond \\
    \label{inteq:79}
    &\leq \delta_2 + 4\delta_1.
\end{align}
This suggests that the error $\delta_1$ arising in the preparation of the canonical purified state is not accumulated by $\mathtt{DMESUB}$.

The above discussion about the channel $\til{\cF}^{\sC\hat{\sA}\hat{\sB}\sB}$ applies similarly to $\til{\cF}_{\rm inv}^{\sC\hat{\sA}\hat{\sB}\sB} = \mathtt{DMESUB}^\dag(\til{\Upsilon}^{\sC_2\sC_3\sA_1\sB_1\sB_2}; \delta_2, t=2)$. Thus, $\til{\cF}_{\rm inv}^{\sC\hat{\sA}\hat{\sB}\sB}$ approximates a unitary channel $(\cU_2^{\sC\hat{\sA}\hat{\sB}\sB})^\dag$ within the error $\delta_2 + 4\delta_1$.

Up to this point, the number of samples of $\rho^{\hat{\sA}}$ and $\sigma^{\hat{\sA}}$ needed for $\til{\cF}^{\sC\hat{\sA}\hat{\sB}\sB}$ and $\til{\cF}_{\rm inv}^{\sC\hat{\sA}\hat{\sB}\sB}$ is
\begin{align}
\label{inteq:99}
    mf 
    &= \til{\cO}\big(1/\delta_2\big)\til{\cO}\Big(\f{d_\sA}{\delta_1}\min\Big\{\f{1}{\rho_{\rm min}^2} + \f{1}{\sigma_{\rm min}^2}, \f{d_\sA^2}{\delta_1^4}\Big\}\Big) \notag\\
    &=\til{\cO}\Big(\f{d_\sA}{\delta_1\delta_2}\min\Big\{\f{1}{\rho_{\rm min}^2} + \f{1}{\sigma_{\rm min}^2}, \f{d_\sA^2}{\delta_1^4}\Big\}\Big).
\end{align}

The rest is almost the same as the discussion in the purified sample access model.
Using $\mathtt{QSVTSIGN}$, we lift up all the non-zero singular values of the top-left block of $-iX^\sC U_2^{\sC\hat{\sA}\hat{\sB}\sB}$ to unity.
Note that the top-left block of $-iX^\sC U_2^{\sC\hat{\sA}\hat{\sB}\sB}$ is given by $\ketbra{\rho_{\rm c}}{\sigma_{\rm c}}^{\hat{\sA}\hat{\sB}} \otimes \sin^{(\rm SV)}(M^{\sB})$, where $M^{\sB} = \tr_{\sA'}\big[\ketbra{\sigma_{\rm c}}{\rho_{\rm c}}^{\sA'\sB}\big]$.


For $\delta_3 \in (0, 1/2)$, let $\cJ^{\sB\sH\hat{\sA}\hat{\sB}} = \mathtt{QSVTSIGN}(\til{\cG}^{\sC\hat{\sA}\hat{\sB}\sB}, \til{\cG}_{\rm inv}^{\sC\hat{\sA}\hat{\sB}\sB}; \delta_3, \beta)$, where $\til{\cG}^{\sC\hat{\sA}\hat{\sB}\sB} = \cX^\sC \circ \til{\cF}^{\sC\hat{\sA}\hat{\sB}\sB}$, $\til{\cG}_{\rm inv}^{\sC\hat{\sA}\hat{\sB}\sB} = \til{\cF}_{\rm inv}^{\sC\hat{\sA}\hat{\sB}\sB}\circ\cX^\sC$, and $X^\sC$ is one-qubit Pauli-$X$ gate.
From Eq.~\eqref{inteq:79}, we see that 
\begin{align}
\label{inteq:78}
    \f{1}{2}\big\|\cJ^{\sB\sH\hat{\sA}\hat{\sB}} - \cU_1^{\sB\sH\hat{\sA}\hat{\sB}}\big\|_\diamond \leq u(\delta_2 + 4\delta_1),
\end{align}
where $u = \cO\big(\log{(1/\delta_3)}/\beta\big)$.
The unitary $\cU_1^{\sB\sH\hat{\sA}\hat{\sB}}$ is a block-encoding of $\ketbra{\rho_{\rm c}}{\sigma_{\rm c}}^{\hat{\sA}\hat{\sB}} \otimes P_\sgn^{(\rm SV)}\big(\sin^{(\rm SV)}(M^\sB)\big)$, and $P_\sgn$ is a polynomial approximation of the sign function.


It should be noted that in the mixed sample access model, the quantum channel $\cP_{\ket{\sigma_{\rm c}}}^{\bC\rarr\hat{\sA}\hat{\sB}}$ cannot be implemented exactly.
To this end, we define a state-preparation channel $\cP_{\til{\sigma}_{\rm c}}^{\bC\rarr\hat{\sA}\hat{\sB}}$ by $\cP_{\til{\sigma}_{\rm c}}^{\bC\rarr\hat{\sA}\hat{\sB}} = \til{\sigma}_{\rm c}^{\hat{\sA}\hat{\sB}}$, and evaluate a distance between $\cP_{\til{\sigma}_{\rm c}}^{\bC\rarr\hat{\sA}\hat{\sB}}$ and $\cP_{\ket{\sigma_{\rm c}}}^{\bC\rarr\hat{\sA}\hat{\sB}}$, obtaining that
\begin{align}
\label{inteq:97}
    \f{1}{2}\big\|\cP_{\til{\sigma}_{\rm c}}^{\bC\rarr\hat{\sA}\hat{\sB}} - \cP_{\ket{\sigma_{\rm c}}}^{\bC\rarr\hat{\sA}\hat{\sB}}\big\|_\diamond 
    &\leq \delta_1,
\end{align}
where we used Eq.~\eqref{inteq:41}.
Thus, we have 
\begin{align}
\label{inteq:100}
    \f{1}{2}\big\|\cJ^{\sB\sH\hat{\sA}\hat{\sB}}\circ\cP_{\til{\sigma}_{\rm c}}^{\bC\rarr\hat{\sA}\hat{\sB}} - \cU_1^{\sB\sH\hat{\sA}\hat{\sB}}\circ\cP_{\ket{\sigma_{\rm c}}}^{\bC\rarr\hat{\sA}\hat{\sB}}\big\|_\diamond
    \leq u\delta_2 + (4u+1)\delta_1,
\end{align}
where we used Eqs.~\eqref{inteq:78} and~\eqref{inteq:97}.


The above discussion applies to both cases where the error is measured in the diamond norm and the fidelity difference.
In the following, we analyze them separately in Secs.~\ref{sec:mixed sample diamond} and~\ref{sec:uhlfidsamp}. These establish Theorem~\ref{thm:Uhlmann mixed state sample}.




%===================================================================================================================================================================================




\subsubsection{Evaluation in the diamond norm}
\label{sec:mixed sample diamond}




When we evaluate the performance of this algorithm in the diamond norm, we set $\beta$ to $\beta_\diamond = 2s_{\rm min}/\pi$. We then apply \Cref{lem:FPAA diamond} and obtain that
\begin{align}
\label{inteq:149}
    \f{1}{2}\big\|\cU_1^{\sB\sH\hat{\sA}\hat{\sB}}\circ\cP_{\ket{0}\ket{\sigma_{\rm c}}}^{\bC\rarr\sH\hat{\sA}\hat{\sB}} - \cU_{\rm ideal}^{\sB\sH\hat{\sA}\hat{\sB}}\circ\cP_{\ket{0}\ket{\sigma_{\rm c}}}^{\bC\rarr\sH\hat{\sA}\hat{\sB}}\big\|_\diamond \leq 3\sqrt{\delta_3},
\end{align}
where $\delta_3 \in (0, 1/2)$, $\cU_{\rm ideal}^{\sB\sH\hat{\sA}\hat{\sB}}$ is a block-encoding of $\ketbra{\rho_{\rm c}}{\sigma_{\rm c}}^{\hat{\sA}\hat{\sB}} \otimes V^\sB$, and $V^\sB$ is the Uhlmann partial isometry from $\ket{\rho_{\rm c}}^{\sA\sB}$ to $\ket{\sigma_{\rm c}}^{\sA\sB}$.
Then, we have that 
\begin{align}
    &\f{1}{2}\big\|\cJ^{\sB\sH\hat{\sA}\hat{\sB}}\circ\cP_{\ket{0}}^{\bC\rarr\sH}\circ\cP_{\til{\sigma}_{\rm c}}^{\bC\rarr\hat{\sA}\hat{\sB}} - \cU_{\rm ideal}^{\sB\sH\hat{\sA}\hat{\sB}}\circ\cP_{\ket{0}\ket{\sigma_{\rm c}}}^{\bC\rarr\sH\hat{\sA}\hat{\sB}}\big\|_\diamond \notag\\
    &\leq \f{1}{2}\big\|\cJ^{\sB\sH\hat{\sA}\hat{\sB}}\circ\cP_{\til{\sigma_{\rm c}}}^{\bC\rarr\sH\hat{\sA}\hat{\sB}} - \cU_1^{\sB\sH\hat{\sA}\hat{\sB}}\circ\cP_{\ket{\sigma_{\rm c}}}^{\bC\rarr\sH\hat{\sA}\hat{\sB}}\big\|_\diamond + \f{1}{2}\big\|\cU_1^{\sB\sH\hat{\sA}\hat{\sB}}\circ\cP_{\ket{0}\ket{\sigma_{\rm c}}}^{\bC\rarr\sH\hat{\sA}\hat{\sB}} - \cU_{\rm ideal}^{\sB\sH\hat{\sA}\hat{\sB}}\circ\cP_{\ket{0}\ket{\sigma_{\rm c}}}^{\bC\rarr\sH\hat{\sA}\hat{\sB}}\big\|_\diamond \\
    \label{inteq:131}
    &\leq  3\sqrt{\delta_3} + u\delta_2 + (4u + 1)\delta_1,
\end{align}
where $u=\cO\big(\log{(1/\delta_3)/\beta_\diamond}\big)$. 
We used Eqs.~\eqref{inteq:100} and~\eqref{inteq:149} in the last line.

Note that the first two terms on the right-hand side of Eq.~\eqref{inteq:131} correspond to the right-hand side of Eq.~\eqref{inteq:103} (with $\delta_1 \gets \delta_3$).
This suggests that, defining a quantum channel $\cJ_\diamond^{\sB\sH\hat{\sA}\hat{\sB}}$ by $\cJ_\diamond^{\sB\sH\hat{\sA}\hat{\sB}} = \mathtt{UhlmannPurifiedSample}(\til{\rho}_{\rm c}^{\hat{\sA}\hat{\sB}}, \til{\sigma}_{\rm c}^{\hat{\sA}\hat{\sB}}; \delta/2, \diamond)$, the parameters should be rescaled by $\delta_2 \gets \delta/(4u)$ and $\delta_3 \gets (\delta/12)^2$ in $\mathtt{UhlmannPurifiedSample}$ (see also Algorithm~\ref{alg:Uhl purif sample}).
Thus, rescaling $\delta_1$ as $\delta_1 = \delta/\big(2(4u+1)\big)$, the overall error is bounded as
\begin{align}
    &\f{1}{2}\big\|\cJ_\diamond^{\sB\sH\hat{\sA}\hat{\sB}}\circ\cP_{\ket{0}}^{\bC\rarr\sH}\circ\cP_{\til{\sigma}_{\rm c}}^{\bC\rarr\hat{\sA}\hat{\sB}} - \cU_{\rm ideal}^{\sB\sH\hat{\sA}\hat{\sB}}\circ\cP_{\ket{0}\ket{\sigma_{\rm c}}}^{\bC\rarr\sH\hat{\sA}\hat{\sB}}\big\|_\diamond \leq \delta.
\end{align}

The number of samples of $\rho^\sA$ and $\sigma^\sA$ is given by
\begin{align}
\label{inteq:45}
    \zeta_\diamond 
    &= umf \\
    &= \cO\Big(\f{1}{\beta_\diamond}\log{(1/\delta_3)}\Big)\til{\cO}\Big(\f{d_\sA}{\delta_1\delta_2}\min\Big\{\f{1}{\rho_{\rm min}^2} + \f{1}{\sigma_{\rm min}^2}, \f{d_\sA^2}{\delta_1^4}\Big\}\Big)    \\
    \label{inteq:44}
    &= \til{\cO}\Big(\f{d_\sA}{\delta^2 s_{\rm min}^3} \min\Big\{\f{1}{\rho_{\rm min}^2} + \f{1}{\sigma_{\rm min}^2}, \f{d_\sA^2}{\delta^4 s_{\rm min}^4}\Big\}\Big).
\end{align}

We use $f\log{d_\sA}$ swap gates to prepare one pair of $\til{\rho}_{\rm c}^{\hat{\sA}\hat{\sB}}$ and $\til{\sigma}_{\rm c}^{\hat{\sA}\hat{\sB}}$. 
Once $w_\diamond = \cO\big(\big(\log{(1/\delta)}\big)^2(\delta s_{\rm min}^2)\big)$ pairs are prepared by $w_\diamond f \log{d_\sA}$ gates, $\mathtt{UhlmannPurifiedSample}$ is performed by $\cO(w_\diamond\log{d_\sA})$ gates. 
Hence, the total number of one- and two-qubit gates is 
\begin{equation}
    \cO\Big(\big(w_\diamond f + w_\diamond\big)\log{d_\sA}\big)\Big) = \cO\big(\zeta_\diamond\log{d_\sA}\big),
\end{equation}
where we used $w_\diamond f = \zeta_\diamond$.
Note that $\cO\big(\log{(d_\sA d_\sB)}\big) = \cO(\log{d_\sA})$, as $d_\sA = d_\sB$.
It suffices to use $\cO(\log{d_\sA})$ qubits at any one time, since this algorithm is performed sequentially.





%==========================================================================================================================================================================


\subsubsection{Evaluation in the fidelity difference}
\label{sec:uhlfidsamp}



Evaluation of the performance of the algorithm in the fidelity difference is almost identical to those provided in Secs.~\ref{sec:uhlfidquery} and~\ref{sec:uhlfidpurifsamp}.

We set $\beta$ to $\beta_\tF = \f{1}{\pi}\max\{2s_{\rm min}, \delta_3/r\}$. When we define $\sqrt{\rF'}$ by 
\begin{align}
    \sqrt{\rF'} = \sqrt{\rF}(U_1^{\sB\sH\hat{\sA}\hat{\sB}}\ket{\rho_{\rm c}}^{\sA\sB}\ket{\sigma_{\rm c}}^{\hat{\sA}\hat{\sB}}\ket{0}^{\sH}, \ket{\sigma_{\rm c}}^{\sA\sB}\ket{\rho_{\rm c}}^{\hat{\sA}\hat{\sB}}\ket{0}^{\sH}),
\end{align}
we have that
\begin{align}
\label{inteq:77}
    &\big|\sqrt{\rF}(\rho^\sA, \sigma^\sA) - \sqrt{\rF'}\big| \leq 2\delta_3.
\end{align}
This follows from the similar calculation from Eq.~\eqref{inteq:104} to Eq.~\eqref{inteq:13} in \Cref{sec:uhlfidquery}.


Let $\cT^{\sB}= \tr_{\sH\hat{\sA}\hat{\sB}}\circ\cJ^{\sB\sH\hat{\sA}\hat{\sB}}\circ\cP_{\ket{0}}^{\bC\rarr\sH}\circ\cP_{\til{\sigma}_{\rm c}}^{\bC\rarr\hat{\sA}\hat{\sB}}$. We obtain that
\begin{align}
    \big|\rF(\rho^\sA, \sigma^\sA) - \rF\big(\cT^{\sB}(\ketbra{\rho_{\rm c}}{\rho_{\rm c}}^{\sA\sB}), \ket{\sigma_{\rm c}}^{\sA\sB}\big)\big| 
    &\leq \f{u}{2}\big\|\til{\cF}^{\sC\hat{\sA}\hat{\sB}\sB'} - \cU_2^{\sC\hat{\sA}\hat{\sB}\sB'}\big\|_\diamond \notag\\
    &\hspace{3pc}+ \f{1}{2}\big\|\cP_{\til{\sigma}_{\rm c}}^{\bC\rarr\hat{\sA}\hat{\sB}} - \cP_{\ket{\sigma_{\rm c}}}^{\bC\rarr\hat{\sA}\hat{\sB}}\big\|_\diamond 
    + \big|\rF(\rho^\sA, \sigma^\sA) - \rF'\big| \\
    \label{inteq:105}
    &\leq u\delta_2 + 4\delta_3 +  (4u+1)\delta_1,
\end{align}
where $u = \cO\big(\log(1/\delta_3)/\beta_\tF\big)$.
Here, we used the fact that $\cJ^{\sB\sH\hat{\sA}\hat{\sB}}$ consists of $u$ use of the channel $\til{\cF}^{\sC\sH\hat{\sA}\hat{\sB}\sB}$ (and $\til{\cF}_{\rm inv}^{\sC\sH\hat{\sA}\hat{\sB}\sB}$) and that $|x-y|\leq 2 |\sqrt{x} - \sqrt{y}|$ for $0 \leq x, y \leq 1$, and also used Eqs.~\eqref{inteq:98},~$\eqref{inteq:97}$, and~\eqref{inteq:77}.

We define a quantum channel $\cJ_\tF^{\sB\sH\hat{\sA}\hat{\sB}}$ by $\cJ_\tF^{\sB\sH\hat{\sA}\hat{\sB}} = \mathtt{UhlmannPurifiedSample}(\til{\rho}_{\rm c}^{\hat{\sA}\hat{\sB}}, \til{\sigma}_{\rm c}^{\hat{\sA}\hat{\sB}}; \delta/2, \tF)$ and note that the parameters are rescaled as $\delta_2 \gets \delta/(4u)$ and $\delta_3 \gets \delta/16$ in $\mathtt{UhlmannPurifiedSample}$ in Algorithm~\ref{alg:Uhl purif sample}, as the first two terms on the right-hand side of Eq.~\eqref{inteq:105} correspond to the right-hand side of Eq.~\eqref{inteq:106}.
Thus, rescaling $\delta_1$ as $\delta_1 = \delta/\big(2(4u+1)\big)$, the overall error is bounded as
\begin{align}
     \rF\big(\cT^{\sB}(\ketbra{\rho_{\rm c}}{\rho_{\rm c}}^{\sA\sB}), \ket{\sigma_{\rm c}}^{\sA\sB}\big) \geq \rF(\rho^\sA, \sigma^\sA) - \delta.
\end{align}

Finally, we evaluate the number of samples of $\rho^{\hat{\sA}}$ and $\sigma^{\hat{\sA}}$ as 
\begin{align}
    \zeta_\tF 
    &= umq \\
    &=\til{\cO}\Big(\f{d_\sA}{\delta^2\beta_\tF^3}\min\Big\{\f{1}{\rho_{\rm min}^2} + \f{1}{\sigma_{\rm min}^2}, \f{d_\sA^2}{\beta_\tF^4\delta^4}\Big\}\Big),
\end{align}
where $mq$ is in Eq.~\eqref{inteq:99}.
The number of one- and two-qubit gates is
$\cO\big(\zeta_\tF \log{d_\sA}\big)$, and $\cO(\log{d_\sA})$ qubits suffice at any one time.







%=======================================================================================================================================================
