

\documentclass[superscriptaddress,11pt,bibliography]{article}


\usepackage[letterpaper,margin=1in,bottom=1in]{geometry}
\usepackage{setspace}
\setstretch{1}
\usepackage{titlesec}
\usepackage{microtype}



\usepackage[page,header]{appendix}


\usepackage[T1]{fontenc}
\usepackage{tgtermes}
\usepackage{newtxtext,newtxmath}
\renewcommand{\rmdefault}{ntx}
\let\Bbbk\relax
\let\openbox\relax
\usepackage{amsmath,amsfonts,amssymb,amstext, amsthm}

\usepackage[font=small,labelfont=bf]{caption}




\usepackage{anyfontsize}


\usepackage[pdftex]{graphicx}


\usepackage[dvipsnames,svgnames]{xcolor}
\usepackage{comment}



\usepackage{mathtools}
\usepackage{nccmath}
\usepackage{blkarray}
\usepackage{here}

\usepackage{array}

\usepackage{afterpage}

\usepackage{makecell}


\usepackage{enumitem}

\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage[most]{tcolorbox}

\usepackage{enumitem}

\usepackage{thmtools}


\usepackage{minitoc}
\usepackage{lipsum}
\usepackage{shorttoc}
\usepackage{etoc}
  


\usepackage{url}
\usepackage[linktocpage, colorlinks]{hyperref}
\hypersetup{
	colorlinks = true,
	citecolor = blue,
	linkcolor = magenta,
	urlcolor = brown,
}

\usepackage[normalem]{ulem}




\usepackage{cleveref}
\Crefname{section}{Sec.}{Secs.}
\Crefname{subsection}{Sec.}{Secs.}
\Crefname{subsection}{Sec.}{Secs.}
\Crefname{figure}{Fig.}{Figs.}
\Crefname{table}{Table}{Tables}
\Crefname{algorithm}{Algorithm}{Algorithms}
\Crefname{theorem}{Theorem}{Theorems}
\Crefname{lemma}{Lemma}{Lemmas}
\Crefname{proposition}{Proposition}{Propositions}
\Crefname{corollary}{Corollary}{Corollaries}


\newtheorem{theorem}{Theorem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{question}[theorem]{Question}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{example}[theorem]{Example}
\newtheorem{conjecture}[theorem]{Conjecture}

\theoremstyle{definition}
\newtheorem*{Proof}{Proof}




\newcommand{\red}[1]{\textcolor{red}{#1}}
\newcommand{\green}[1]{\textcolor{green}{#1}}
\newcommand{\blue}[1]{\textcolor{blue}{#1}}
\newcommand{\cyan}[1]{\textcolor{cyan}{#1}}
\newcommand{\magenta}[1]{\textcolor{magenta}{#1}}


\newcommand{\ket}[1]{|#1\rangle}
\newcommand{\bra}[1]{\langle#1|}
\newcommand{\ketbra}[2]{|#1\rangle\langle#2|}
\newcommand{\braket}[2]{\langle #1 \vert #2 \rangle}

\DeclareMathOperator{\tr}{Tr}

\newcommand{\rid}{{\mathrm{id}}}
\newcommand{\bR}{\mathbb{R}}
\newcommand{\bC}{\mathbb{C}}
\newcommand{\bN}{\mathbb{N}}
\newcommand{\bZ}{\mathbb{Z}}
\newcommand{\bI}{\mathbb{I}}
\newcommand{\bE}{\mathbb{E}}

\newcommand{\cA}{{\mathcal{A}}}
\newcommand{\cB}{{\mathcal{B}}}
\newcommand{\cC}{{\mathcal{C}}}
\newcommand{\cD}{{\mathcal{D}}}
\newcommand{\cE}{{\mathcal{E}}}
\newcommand{\cF}{{\mathcal{F}}}
\newcommand{\cG}{{\mathcal{G}}}
\newcommand{\cH}{{\mathcal{H}}}
\newcommand{\cI}{{\mathcal{I}}}
\newcommand{\cJ}{{\mathcal{J}}}
\newcommand{\cK}{{\mathcal{K}}}
\newcommand{\cL}{{\mathcal{L}}}
\newcommand{\cM}{{\mathcal{M}}}
\newcommand{\cN}{{\mathcal{N}}}
\newcommand{\cO}{{\mathcal{O}}}
\newcommand{\cP}{{\mathcal{P}}}
\newcommand{\cQ}{{\mathcal{Q}}}
\newcommand{\cR}{{\mathcal{R}}}
\newcommand{\cS}{{\mathcal{S}}}
\newcommand{\cT}{{\mathcal{T}}}
\newcommand{\cU}{{\mathcal{U}}}
\newcommand{\cV}{{\mathcal{V}}}
\newcommand{\cW}{{\mathcal{W}}}
\newcommand{\cX}{{\mathcal{X}}}
\newcommand{\cY}{{\mathcal{Y}}}
\newcommand{\cZ}{{\mathcal{Z}}}

\newcommand{\rA}{{\mathrm{A}}}
\newcommand{\rB}{{\mathrm{B}}}
\newcommand{\rC}{{\mathrm{C}}}
\newcommand{\rD}{{\mathrm{D}}}
\newcommand{\rE}{{\mathrm{E}}}
\newcommand{\rF}{{\mathrm{F}}}
\newcommand{\rG}{{\mathrm{G}}}
\newcommand{\rH}{{\mathrm{H}}}
\newcommand{\rI}{{\mathrm{I}}}
\newcommand{\rJ}{{\mathrm{J}}}
\newcommand{\rK}{{\mathrm{K}}}
\newcommand{\rL}{{\mathrm{L}}}
\newcommand{\rM}{{\mathrm{M}}}
\newcommand{\rN}{{\mathrm{N}}}
\newcommand{\rO}{{\mathrm{O}}}
\newcommand{\rP}{{\mathrm{P}}}
\newcommand{\rQ}{{\mathrm{Q}}}
\newcommand{\rR}{{\mathrm{R}}}
\newcommand{\rS}{{\mathrm{S}}}
\newcommand{\rT}{{\mathrm{T}}}
\newcommand{\rU}{{\mathrm{U}}}
\newcommand{\rV}{{\mathrm{V}}}
\newcommand{\rW}{{\mathrm{W}}}
\newcommand{\rX}{{\mathrm{X}}}
\newcommand{\rY}{{\mathrm{Y}}}
\newcommand{\rZ}{{\mathrm{Z}}}

\newcommand{\sA}{{\mathsf{A}}}
\newcommand{\sB}{{\mathsf{B}}}
\newcommand{\sC}{{\mathsf{C}}}
\newcommand{\sD}{{\mathsf{D}}}
\newcommand{\sE}{{\mathsf{E}}}
\newcommand{\sF}{{\mathsf{F}}}
\newcommand{\sG}{{\mathsf{G}}}
\newcommand{\sH}{{\mathsf{H}}}
\newcommand{\sI}{{\mathsf{I}}}
\newcommand{\sJ}{{\mathsf{J}}}
\newcommand{\sK}{{\mathsf{K}}}
\newcommand{\sL}{{\mathsf{L}}}
\newcommand{\sM}{{\mathsf{M}}}
\newcommand{\sN}{{\mathsf{N}}}
\newcommand{\sO}{{\mathsf{O}}}
\newcommand{\sP}{{\mathsf{P}}}
\newcommand{\sQ}{{\mathsf{Q}}}
\newcommand{\sR}{{\mathsf{R}}}
\newcommand{\sS}{{\mathsf{S}}}
\newcommand{\sT}{{\mathsf{T}}}
\newcommand{\sU}{{\mathsf{U}}}
\newcommand{\sV}{{\mathsf{V}}}
\newcommand{\sW}{{\mathsf{W}}}
\newcommand{\sX}{{\mathsf{X}}}
\newcommand{\sY}{{\mathsf{Y}}}
\newcommand{\sZ}{{\mathsf{Z}}}

\newcommand{\sIn}{\mathsf{In}}
\newcommand{\sOut}{\mathsf{Out}}



\newcommand{\tF}{\mathtt{F}}



\newcommand{\bs}[1]{\boldsymbol{#1}}
\newcommand{\mbb}[1]{\mathbb{#1}}
\newcommand{\EH}{\mbb{E}_{U \sim {\sf H}}}

\newcommand{\f}{\frac}
\newcommand{\dagg}{\dagger}

\newcommand{\rarr}{\rightarrow}
\newcommand{\larr}{\leftarrow}
\newcommand{\vphi}{\varphi}
\newcommand{\til}{\tilde}

\newcommand{\supp}{\mathrm{supp}}
\newcommand{\ctrl}{\mathrm{ctrl}}
\newcommand{\ad}{\mathrm{ad}}
\newcommand{\poly}{\mathrm{poly}}
%\newcommand{\sign}{\mathrm{sign}}
\newcommand{\sgn}{\mathrm{sgn}}
\newcommand{\im}{\mathrm{im}}


\renewcommand{\thefootnote}{\fnsymbol{footnote}}


\newtcolorbox{dashedbox}[1][]{
  enhanced,
  sharp corners,
  boxrule=0.5pt,
  colback=white,
  colframe=black,
  fonttitle=\bfseries,
  title=#1,
  dash pattern=on 2pt off 2pt,
  borderline={0.5pt}{0pt}{black!60!white,dashed},
  before skip=5pt, after skip=5pt,
  boxsep=5pt
}



\begin{document}


\pagestyle{empty}

\begin{titlepage}


%======================================================================================================================================================================================================

\begin{flushright}
YITP-25-135
\end{flushright}

\vspace*{0.5cm}


%======================================================================================================================================================================================================


\begin{center}
{\LARGE\bfseries Quantum algorithms for Uhlmann transformation \par}


%======================================================================================================================================================================================================

\vspace*{0.4cm}

{\large
Takeru Utsumi$^{1, }$\footnote{takeru-utsumi@g.ecc.u-tokyo.ac.jp},
Yoshifumi Nakata$^{2, }$\footnote{yoshifumi.nakata@yukawa.kyoto-u.ac.jp},
Qisheng Wang$^{3, }$\footnote{QishengWang1994@gmail.com}, and 
Ryuji Takagi$^{1, }$\footnote{ryujitakagi.pat@gmail.com}
}

\vspace{0.3cm}

{\small
$^{1}$Graduate School of Arts and Sciences, The University of Tokyo, 3-8-1 Komaba, Meguro-ku, Tokyo, Japan\\
$^{2}$Yukawa Institute for Theoretical Physics, Kyoto University, Oiwake-cho, Kitashirakawa, Sakyo-ku, Kyoto, Japan\\
$^{3}$School of Informatics, University of Edinburgh, EH8 9AB Edinburgh, United Kingdom
}



\vspace{0.2cm}


\today


\vspace*{-1cm}



%======================================================================================================================================================================================================



\vspace*{1.5cm}


\begin{abstract}

Uhlmann's theorem is a central result in quantum information theory, which associates the closeness of two quantum states with that of their purifications. The theorem also well characterizes a fundamental task: how close a pure quantum state can be transformed into another state via local operations acting only on its subsystem. The optimal transformation for this task is called the Uhlmann transformation, which has broad applications in various information-processing tasks. However, its quantum circuit implementation and computational cost have remained unclear, limiting the utility of the transformation.

In this work, we fill this gap by proposing quantum algorithms that realize the Uhlmann transformation in query and sample access models. Notably, our Uhlmann transformation algorithms can be polynomial-time for low-rank states, exhibiting an \textit{exponential} improvement over the approach of \hyperlink{cite.metger2023stateqipspace}{Metger and Yuen (FOCS 2023)} and other naive approaches based on quantum state tomography. In addition, we derive a lower bound on the query and sample complexities of the Uhlmann transformation for a deeper understanding of its algorithmic features.

We apply our Uhlmann transformation algorithms to fidelity estimation between two states, and substantially improve the previous best query and sample complexities by \hyperlink{cite.gilyen2022improvedfidelity}{Gily\'en and Poremba (TQC 2022)} and \hyperlink{cite.Liu2024geometricmean}{Liu, Wang, Wilde, and Zhang (\textit{npj Quantum Inf.}\ 2025)}. Specifically, for the estimation with additive error $\delta$, our approaches yield query complexity $\tilde O(\min\{\frac{\kappa}{\delta}, \frac{r}{\delta^2}\})$ and sample complexity $\tilde O(\min\{\frac{\kappa^2}{\delta^2}, \frac{r^2}{\delta^4}\})$, compared to the previous results of $\tilde O(\min\{\frac{\kappa^4}{\delta}, \frac{r^{2.5}}{\delta^{5}}\})$ queries and $\tilde O(\min\{\frac{\kappa^9}{\delta^3}, \frac{r^{5.5}}{\delta^{12}}\})$ samples, where $\kappa$ is the reciprocal of the minimum non-zero eigenvalue of the states and $r$ is their rank. 

We further discuss other applications to several information-theoretic tasks, including entanglement transmission, quantum state merging, and the algorithmic implementation of the Petz recovery map, providing a comprehensive evaluation of their computational costs. These results, hence, contribute to the practical realization of such widely recognized and useful protocols.


\end{abstract}
\end{center}
\end{titlepage}


\tableofcontents


\clearpage
\setcounter{page}{1}  
\pagenumbering{arabic} 
\pagestyle{plain}


\section{Introduction}
\label{sec:introduction}
\input{0_introduction}



\section{Main results}
\label{sec:mainresults}
\input{1_mainresults}

\section{Summary and outlooks}
\label{sec:conclusion}
\input{5_conclusion}




\section{Preliminaries}
\label{sec:preliminaries}
\input{2_preliminaries}


\section{Uhlmann transformation algorithm in the purified query access model}
\label{sec:purif query}
\input{3_1_purified_query}
\section{Uhlmann transformation algorithm in the purified sample access model}
\label{sec:purif sample}
\input{3_2_purified_sample}
\section{Uhlmann transformation algorithm in the mixed sample access model}
\label{sec:mixed sample}
\input{3_3_mixed_sample}


\section{Application 1: Square root fidelity estimation with the Uhlmann transformation algorithm}
\label{sec:application}
\input{4_1_application}

\section{Application 2: Decoupling approach with the Uhlmann transformation algorithm}
\label{sec:decoupling and uhlmann}
\input{4_2_application}

\section{Application 3: Petz recovery map with the Uhlmann transformation algorithm}
\label{sec:sample petz}
\input{4_3_application}


%======================================================================================================================================================================================================



\section*{Acknowledgments}
\addcontentsline{toc}{section}{Acknowledgments}
We thank Kosuke Mitarai, Keisuke Fujii, Takaya Matsuura, Zhaoyi Li, and Seth Lloyd for helpful discussions. We are also grateful to Hayato Arai for facilitating our collaboration.

T.U.\ acknowledges the support of JST CREST Grant Number JPMJCR23I3 and JST SPRING Grant Number JPMJSP2108.
Y.N.\ is supported by MEXT KAKENHI Grant-in-Aid for Transformative Research Areas (A) ``Extreme Universe'' Grant Numbers JP21H05182 and JP21H05183, JST CREST Grant Number JPMJCR23I3, and JST PRESTO Grant Number JPMJPR2456.
Q.W.\ acknowledges the support of the Engineering and Physical Sciences Research Council under Grant \mbox{EP/X026167/1}.
R.T.\ acknowledges the support of JST CREST Grant Number JPMJCR23I3, JSPS KAKENHI Grant Number JP24K16975, JP25K00924, and MEXT KAKENHI Grant-in-Aid for Transformative
Research Areas A ``Extreme Universe” Grant Number JP24H00943.



\bibliographystyle{alpha_mod}
\addcontentsline{toc}{section}{References}
\bibliography{ref}


%======================================================================================================================================================================================================




\appendix
\part*{\LARGE Appendix}
\addcontentsline{toc}{section}{Appendix}



\crefname{appendix}{appendix}{appendices}
\Crefname{appendix}{Appendix}{Appendices}


\crefalias{section}{appendix}
\crefalias{subsection}{appendix}
\crefalias{subsubsection}{appendix}


\input{appendix1}
\input{appendix2}






\end{document}
